---
title: "Indonesian Shrimp Export to United States of America: Trends, Performance, and Competitiveness."
author: "Hariono"
date: "4/29/2020"
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
###Instal Usefull Package, For the first time running the code may be you need to install packages below, to run just delete the "#"
#install.packages("kableExtra")
#install.packages("tidyverse")
#install.packages("dplyr")

###Load Usefull Package
library(kableExtra) ###Make a good table
library(tidyverse)
library(dplyr)
library(ggplot2)

```

```{r import data, include=FALSE, echo=FALSE, warning=FALSE}
###Import Usefull Data
Indo_Import <- read_csv("https://raw.githubusercontent.com/Harionotph45/Final-Project-RCA-/master/Indo%20Import%20From%20US.csv") #This is data Indonesia Import from USA
Indo_USVal <- read_csv("https://raw.githubusercontent.com/Harionotph45/Final-Project-RCA-/master/Indo%20to%20US%20Value.csv") #This is data Indonesia Export to USA (Value)
Indo_USVol <- read_csv("https://raw.githubusercontent.com/Harionotph45/Final-Project-RCA-/master/Indo%20to%20US%20Volume.csv") #This is data Indonesia Export to USA (Volume)
World_to_US <- read_csv("https://raw.githubusercontent.com/Harionotph45/Final-Project-RCA-/master/World%20to%20US.csv") #This is data World Export to USA (Value)
Ref_HS <- read_csv("https://raw.githubusercontent.com/Harionotph45/Final-Project-RCA-/master/Ref%20HS%20Code.csv") #This is a reference of product code, so we know the Categories(Fish or not) and The Group Commodities (Ex: Shrimp, Tuna, Mollusk, etc), and the forms (ex: Frozen, Fresh, Prepared, etc)

####NOTE : All of the value in the data is in thousands USD, and all of the volume is in Tons

#Select Ref_HSCode column as needed
Ref_HSCode <- Ref_HS %>% select("HS Code 6","Group_Commodities","Commodities","Form") ###The original Ref_HS data contains columns named in Indonesian language Which is we don't need in this analysis

```

```{r Customize The Analysis, include=FALSE, echo=FALSE, warning=FALSE}

####Preparing data to check the avaibility of Group Commodities and Years to Analize
###Join data Indonesia Export to US (Left Join, it will make more specific because Reference HS Code contains all goods exported to Indonesia not only to US) with HS Code and filtered only Fish Categories

Indo_USVal_Check <- Indo_USVal %>% pivot_longer(cols = c(-`Product code`,-`Product label`,-`Categories`), names_to = "Year", 
                                              values_to = "Value") %>%  left_join(Ref_HSCode,by = c("Product code"="HS Code 6"))%>% filter(Categories=="Fish")


##### The Code allowed You to Costumize which Commodities and What Years You want to analyze

##Customize the commodity to analyze
a <- "Shrimp" ####This "a" is a group commodity name, you can change it to another commodity, to know what group commodities are available you can check with the code below :
unique(Indo_USVal_Check$`Group_Commodities`)###This code will check the names of all group commodities, just choose and copy paste it and change the the definition of variable "a" above

##Customize the range of year to analyze 
yi <- 2008    ####This yi is the start of year you will analyze, you can change it as long as the data available
yx <- 2018    ####This yx is the end of year you will analyze, you can change it so you don't need to worry if there is an update to the data 

###To check what years are available to analyze you can run the code below :
unique(Indo_USVal_Check$`Year`) ###You should aware that this will check only the year in data Indonesia export to US (Value), if you want to add a new updated year, please make sure you update all of the data, because some analysis needs more than one dataframe to run


####NOTE : Please Bear in mind that the customize only works for figures and tables (the data presented, the title and the label will be changed based on your customization), but the manual analysis/arguments (which is typed manually including the title of this paper) and the evidences to support the arguments will not change, So you still need to  work on that.

```

### INTRODUCTION

Fisheries are one of the most important sectors that supports more than 200 million livelihoods who work in harvesting, processing, marketing, distribution, and supply industries (FAO,2008). Fish is also a source of protein for humans, providing ~20% of the animal protein consumed by more than 1.5 billion people in the world (FAO, 2008). In the last three decades, world food fish production from aquaculture has increased by almost 12 times, at an average annual rate of 8.8% (FAO, 2012).  Asia accounts for more than 87% of the world's total fish production from aquaculture (FAO, 2012). It is About 38% of the fish and fishery produce are traded internationally(Anon, 2008). About half of the trade-in fish and fishery products originated from developing countries and are destined for the developed markets of Europe, the United States of America, and Japan (Anon, 2008). 

As an archipelago, Indonesia has great potential in the fisheries industry, mainly because it has a reasonably broad sea area, 7.9 million km2, and coastline along the 80,791 km2 with extensive aquaculture and fish pond spread in some parts of the country. Indonesia has a vast 81 percent water area of the total area, so Indonesia may have much potential to rule the business world fisheries. Besides tuna, Indonesia is one of the world's largest shrimp exporting countries, especially to the United States, Japan, and EU markets (Wati, 2013). This study is an attempt to analyze the trend, performance, and competitiveness of Indonesian shrimp exports to the United States market.


### MATERIAL AND METHODS

The export and import data were collected from the United Nations UN COMTRADE statistical database. All of the values have been referred in US dollars, and the volumes were in tons. Classification of HS Code to group commodities was based on the data from the Ministry of Marine Affairs and Fisheries. 

The trade balance is the difference between Indonesia's exports to the USA and Indonesia's imports from the USA. It was calculated to analyze the bilateral trade patterns that occur between Indonesia and the United States for shrimp-based products.  The results would suggest either Indonesia is the major exporter or importer of shrimp to the United States.

Revealed Comparative Advantage (RCA),  one of the most popular methods for studying the comparative advantage of exports, was used in this study. RCA uses the trade pattern to identify the sectors in which an economy has a comparative advantage by comparing the country of interests' trade profile with the world average. The RCA index is defined as the ratio of two shares. The numerator is the share of a country's total exports of the commodity of interest in its total exports. The denominator is share of world exports of the same commodity in total world exports. According to Balassa (1965), the formula to calculate the RCA is as follows:


**RCA =  (XJi/XJt )/(XWi/XWt )**
Where; 
XJi =  _Country J’s exports of good i;_
XJt	= _Country J’s total exports;_
XWi	= _Total world trade in good i;_
XWt	= _Total World Trade._ 

if RCA>1, then the good has a comparative advantage, accelerating product for export-oriented. But, if RCA < 1, the country is said to have a relative disadvantage in the good or industry. 


### RESULTS


```{r create usefull function, include=FALSE, echo=FALSE, warning=FALSE}

#Create Usefull Function
##Create Function to Format number to Percent
###Create Percentage Function, we use percent in some of analysis so make function to calculate percent is necessary
percent <- function(x, digits = 2, 
                    format = "f", ...) 
  {paste0(formatC(100 * x, format = format, digits = digits, ...), "%")
} # use the function by :  percent(x)

###Create Function to Reduce Digits Number, the value in data is in thousands but it still very long digits, to make it easier to read lets make it in million
million <- function(x){as.numeric(as.character((format(round(x / 1e3, 0),format = format))))
  }###NOTE: this function just reduce 3 digits because the current data is already in 000 USD, Use the function by : million()

```

```{r prepared export all fish to usa in yx, include=FALSE, echo=FALSE, warning=FALSE}

#We want to analyze what commodities exported from Indonesia to USA in the yx (year x, "look at chunk customize the Analysis" above to know the detail)
### Because the data Indo_USVal doesn't have column categories (to filter only fish) and column group commodities, we need to look up at Ref_HSCode. So we left join the data 

#Join Indo_USVal with HS Ref
Indo_USVal_ref <- Indo_USVal %>% pivot_longer(cols = c(-`Product code`,-`Product label`,-`Categories`), ####We use pivot longer to make the data (year) possible to filter
                                              names_to = "Year", 
                                              values_to = "Value") %>%
  left_join(Ref_HSCode,by = c("Product code"="HS Code 6"))

#Now we only need column : Group Commodities, Value, and Percentage (will be used as inline code)   
###Group and summarize data
Indo_to_USVal <- Indo_USVal_ref %>% 
  filter(Year== yx, Categories=="Fish") %>% ####Filtered by year and categories as explained before
  group_by(Group_Commodities) %>%
  summarise(Value_000USD = million(sum(Value)))%>% 
  arrange(desc(Value_000USD)) %>%
  mutate(Percentage = percent(Value_000USD/sum(Value_000USD)))%>%
    mutate(Value_M = paste(Value_000USD,"Million")) ###This new column made just to be used for inline code,not in the plot

####Calculate value of each commodities for inline codes
TotalValue <- Indo_to_USVal %>% pull(Value_000USD) %>% sum()%>%paste(.,"Million") ###use paste to add needed information
ValueShrimp <- Indo_to_USVal %>% filter(Group_Commodities=="Shrimp")%>% pull(Value_M)
ValueCrabs <- Indo_to_USVal %>% filter(Group_Commodities=="Crabs" )%>% pull(Value_M)
ValueTuna <- Indo_to_USVal %>% filter(Group_Commodities=="Tuna & Tuna Like Species")%>% pull(Value_M)
ValueTilapia <- Indo_to_USVal %>% filter(Group_Commodities=="Tilapias" )%>% pull(Value_M)
ValueMollusk <- Indo_to_USVal %>% filter(Group_Commodities=="Mollusk" )%>% pull(Value_M)
ValueOtherFish <- Indo_to_USVal %>% filter(Group_Commodities=="Other Fish"   )%>% pull(Value_M)

####Calculate Percentage of each commodities for inline codes
PercentShrimp <- Indo_to_USVal %>% filter(Group_Commodities=="Shrimp")%>% pull(Percentage) 
PercentCrabs <- Indo_to_USVal %>% filter(Group_Commodities=="Crabs" )%>% pull(Percentage)
PercentTuna <- Indo_to_USVal %>% filter(Group_Commodities=="Tuna & Tuna Like Species")%>% pull(Percentage)
PercentTilapia <- Indo_to_USVal %>% filter(Group_Commodities=="Tilapias" )%>% pull(Percentage)
PercentMollusk <- Indo_to_USVal %>% filter(Group_Commodities=="Mollusk" )%>% pull(Percentage)
PercentOtherFish <- Indo_to_USVal %>% filter(Group_Commodities=="Other Fish"   )%>% pull(Percentage)
```

In `r yx`, shrimp was accounted for as the highest export value to the United States among the other commodities, which is shown in  Figure 1. The total value of Indonesian's fisheries products exported to the United States was `r TotalValue` USD, while  `r PercentShrimp` of them or `r ValueShrimp` USD was shrimp, and followed by crabs by `r PercentCrabs` or `r ValueCrabs` USD, Tuna & Tuna like species by `r PercentTuna` or `r ValueTuna` USD, Tilapia `r PercentTilapia` or `r ValueTuna` USD, Mollusk `r PercentMollusk` or `r ValueTuna` USD, and Other fish by `r PercentOtherFish` or `r ValueOtherFish` USD.

```{r fig.1 export to usa in yx, echo=FALSE}
###Plotting Composition of fish
ggplot(Indo_to_USVal,aes(x=Group_Commodities,y=Value_000USD))+
  geom_bar(stat="identity", fill="steelblue")+ labs(title= paste("Fig.1. Indonesian Fish Exported to USA \nBased on Group Commodities in ",yx),x="Group Commodities", y = "Value (Million US$)", color="") ###Use /n to wrap the title line, as the original line is very long, and use paste to combine variable and text, so that when commodities or year changed, the titlle would be changed too
```


```{r prepare fig2 trend volval, include=FALSE, echo=FALSE, warning=FALSE}
##In this second figure we want to know the trend of group commodities "a" during year i to year x by volume and value, so first we need to join the data that contains value (Indo_USVal) and the data that contains volume (Indo_USVol) 

##Before join those two tables, we need to make sure they have the same format, So we need to pivot longer the data Indo_USVol
Indo_USDVol_longer <- Indo_USVol %>% pivot_longer(cols = c(-`Product code`,-`Product label`),
                                        names_to = "Year", 
                                        values_to = "Volume") 

###Join data Volume and Value export to USA
Indo_VOlVal <- Indo_USVal_ref %>% 
  left_join(select(Indo_USDVol_longer,-`Product label`),
            by = c("Product code"="Product code", "Year"="Year"))####Joining by based on two columns that should be the same on both dataframes (Product Code and Year)
  
```

```{r composition forms, include=FALSE, echo=FALSE, warning=FALSE}
##In this analysis we want to know what are forms (ex: frozen, dried, fresh, etc) of product a exported in the year x 
## To do this we need to filter the data by commodites a and year x, and then make summarise that grouped by forms
### Filter and Group the Data
Indo_to_USVal_Form <- Indo_USVal_ref %>% 
  filter(Year== yx, Categories=="Fish", Group_Commodities==a) %>%
  group_by(Form) %>%
  summarise(Value_000USD = million(sum(Value))) %>% 
  arrange(desc(Value_000USD)) %>%
  mutate(Proportion = Value_000USD/sum(Value_000USD))%>% ####To calculate the proportion of each form
  mutate(Value_M = paste(Value_000USD,"Million")) ###add this to custumize the value and make it easier to use for inline code

###Create variables needed for inline Code
ValueFrozen <- Indo_to_USVal_Form %>%filter(Form=="Frozen") %>% pull(Value_M)
ValuePrepared <- Indo_to_USVal_Form %>% filter(Form=="Prepared Or Preserved") %>% pull(Value_M)
PercentFrozen <- Indo_to_USVal_Form %>% filter(Form=="Frozen") %>% pull(Proportion)%>%percent()
PercentPrepared <- Indo_to_USVal_Form %>% filter(Form=="Prepared Or Preserved") %>% pull(Proportion)%>%percent()
```


Figure 2. shows that in general, shrimp exports from Indonesia to the United States have increased in the last ten years (2008-2018) both in volume and value. However, the increase is more in terms of value than the quantity. Based on the further analysis, it is also suggested that the exported shrimp products was mostly frozen for about `r PercentFrozen` or `r ValueFrozen` USD, followed by prepared or preserved with proportion `r PercentPrepared` or `r ValuePrepared` USD, while other forms such as live, smoked, and dried are each less than 1%.


```{r fig2 trend volval, echo=FALSE}
###We already prepared the data in chunk prepare fig2 above, Now we make a Plot from that dataframe
###To plot trend of value and volume of commodities a during year i to year x we use ggplot
Indo_VOlVal %>% filter(Group_Commodities==a, Year %in% (yi:yx))%>% #### this "%n%" means in between
  select(Year,Value,Volume) %>% group_by(Year)%>% 
  summarise(Value_USD = million(sum(Value)), Volume_Tons=million(sum(Volume))) %>%
  ggplot(aes(x=Year, group=1))+geom_line(aes(y=Value_USD, colour = "Value (000 USD)"))+
  geom_line(aes(y=Volume_Tons, colour = "Volume (Tons)"))+ ###geom_line two times because we have two data two plot which are value and volume
  labs(title= paste("Fig.2. Performance of Indonesian",a,"Export \nto USA during ",yi,"-",yx),x="Year", y = "Fish Exports;\n Value (Million $US), Volume (Tons)", color="") + scale_color_manual(labels = c(paste(a," export to USA (value)"), paste(a," export to USA (volume)")), values = c("blue", "red"))+ theme(legend.position="top")
```

The trade balance shown in Table 1. revealed that net trade for shrimp-based fishery products exported from Indonesia to the United States was positive. The number indicates that even though there is a bilateral trade between Indonesia and the United States in shrimp products, Indonesia is a major exporter of shrimp to the United States rather than being an importer of shrimp from the United States.

```{r prepared table1. trade balance, include=FALSE, echo=FALSE, warning=FALSE}
###In this analysis, we want to create a table of trade balance, which will consist columns : Year, Export, Import, and Trade balance(Export-Import) 
###Because data export and import in the separate dataframe, so we need to join it first
#Before join table, we have to make sure it has the same columns, hence we Pivot Longer Data Import
Indo_Import_longer <- Indo_Import %>% 
  pivot_longer(cols = c(-`Product code`,-`Product label`),
               names_to = "Year", values_to = "Import (US$") 

#Join Data Export & Import
Indo_ExIm <- Indo_USVal_ref %>% 
  left_join(select(Indo_Import_longer,-`Product label`),
            by = c("Product code"="Product code", "Year"="Year"))

#Calculated Trade Balance 
Indo_Balance <- Indo_ExIm %>% 
  filter(Group_Commodities== a, Year %in% (yi:yx), Categories=="Fish") %>%
  group_by(Year) %>%
  summarise(Indo_Export_US = sum(Value),
            Indo_Import_US = sum(`Import (US$`))%>% 
  mutate("Trade Balance" = Indo_Export_US - Indo_Import_US)
```

```{r table1 trade balance, echo=FALSE}
###We use kable to make a good table
kable(Indo_Balance, booktabs = TRUE, caption =  paste("Trade balance of ",a," between Indonesia and USA during ",yi,"-",yx), col.names = c("Year","Indonesia export (000 USD)", "Indonesia Import (000 USD)","Trade Balance (000 USD)"), escape = FALSE) %>%
  kable_styling(font_size = 14, latex_options = "scale_down") 
```

```{r prepared fig. 3 RCA, include=FALSE, echo=FALSE, warning=FALSE}
####In this last analysis we want to calculate RCA on each period (year)
####The RCA calculation need 4 variable : total indonesian export to USA (commodities a), total indonesian export to USA (all goods), Total World export to US (commodities a), Total World export to US (commodities a)
###First, we calculated all of those variables needed one by one

###1Total Indonesia Export to US (Commodities a)
Indo_a <- Indo_ExIm %>% filter(Group_Commodities == a) %>% 
  group_by(Year) %>% summarise(Value = sum(Value))

###2Total Indonesia Export to US (All Goods)
Indo_all <- Indo_USVal_ref %>% filter(Categories=="All Goods") %>%
  select(Year, Value)
###3Total World Export to US (All Goods)
World_all <- World_to_US %>% 
  pivot_longer(cols = c(-`Code`,-`Product label`,-`Categories`),
               names_to = "Year",
               values_to = "Value") %>% 
  filter(Code=="All Goods") %>% select(Year, Value)

###4Total World Export to US (Commodities a)
World_a <- World_to_US %>% pivot_longer(cols = c(-`Code`,-`Product label`,-`Categories`),
                                        names_to = "Year", 
                                        values_to = "Value") %>%
  left_join(Ref_HSCode,by = c("Code"="HS Code 6")) %>%
    filter(Group_Commodities == a) %>% select(Year, Value) %>%
  group_by(Year) %>% summarise(Value = sum(Value))
  

###Calculate the RCA
##Join all of the those 4 variables needed to one dataframe
Indo_RCA <- Indo_a %>% 
  left_join(Indo_all, by = c("Year"="Year")) %>%
  left_join(World_a, by = c("Year"="Year")) %>%
  left_join(World_all, by = c("Year"="Year")) %>% filter(Year %in% (yi:yx))%>%
  mutate(RCA = round((Value.x / Value.y)/(Value.x.x / Value.y.y)),2)

###Creates variables for inline code
AverageRCA <- Indo_RCA %>% pull(RCA) %>% mean() %>% round(.,2)
RCA2010 <- Indo_RCA %>% filter(Year==2010) %>% pull(RCA)
RCA2018 <- Indo_RCA %>% filter(Year==2018) %>% pull(RCA)

```
The RCA of Indonesian shrimp exports to the United States are presented in Figure 3 revealed that it was highly positive (RCA> 1) with an average `r AverageRCA`. This shows that Indonesia has a strong comparative advantage in exporting shrimp to the United States. Moreover, the figure also explains that generally, in the last ten years, despite a gradual decrease from 2008 to 2010, the RCA  has increased from `r RCA2010` in 2010 to `r RCA2018` in 2018.

```{r fig. 3 RCA, echo=FALSE}
###Plotting RCA
ggplot(Indo_RCA,aes(x=Year,y=RCA))+ geom_point()+geom_text(aes(label=RCA),hjust=-0.5, vjust=1)+
  labs(x="Year", y = "Revealed Comparative Advantage (RCA)") +ggtitle(paste("Fig.4. Revealed Comparative Advantage (RCA) of \n Indonesian's",a,"exports to USA during ",yi,"-",yx))
```

The best period is from 2010 to 2014 when the RCA continuously increased, while after 2014, it becomes volatile, which tends to decline. The decrease in RCA could be caused by many factors such as the decrease in the volume of demand for frozen shrimp by the importing country, the rejection of Indonesian shrimp by the importing country as a result of antibiotic use, the high or low value of a country's GDP, and the limitations and rules set by the importing country (Mufa'ah, 2016). But mainly in the case of shrimp, Saputri (2017) suggest that the primary factor is the presence of large shrimp companies or suppliers from competing countries such as Vietnam and India. Those countries have become the largest shrimp exporter in the world, which makes the competition very tight and an obstacle for Indonesia. On the other hand, Indonesia still lacks in the application of production technology. 

Globalization makes the market competition very hard these days. Free trade encourages every country in the world to take appropriate and effective steps to improve its competitiveness. Indonesia, which has a lot of natural resources and abundant workforce with competitive wages as supporting factors for competitiveness, turns out to be insufficient to create a competitive advantage. To improve competitiveness, Indonesia needs to conduct various evaluations and efforts so as not to lag behind competing countries.


### CONCLUSION AND SUGGESTION

Shrimp has the highest value among other fish commodities exported to the United States in `r yx`. Most of the shrimp products were shipped in frozen and prepared or preserved. The positive trade balance also suggests that in the bilateral relations, Indonesia is a major exporter of shrimp to the United States. Furthermore, it is revealed that shrimp is a highly competitive export commodity as the value of RCA is far greater than one. However, in the last five years, started from 2014, the RCA's trend has fluctuated and tends to decrease due to a large amount of competition from shrimp suppliers from other countries such as Vietnam and India.

One of the best ways to increase the competitiveness of Indonesian shrimp commodities is by improving the quality and quantity of shrimp. The provision of knowledge about intensive maintenance and handling to shrimp farmers and producers is needed, considering that shrimp is a perishable commodity. Besides, shrimp farmers or producers are also expected to be able to optimize value-added commodities.


### REFERENCES

Anon. (2008). Fisheries and International Trade- Department for International Development (DFID) and
    Marine Resources Assessment Group (MRAG)- Policy Brief 7, United Kingdom

Balassa, Bela. (1965). "Trade Liberalisation and 'Revealed' Comparative Advantage". The
    Manchester School of Economic and Social Studies, Vol. 33, Manchester, 1965, pp. 99-123.

Food and Agriculture Organization (FAO). (2008). Climate change implications for fisheries and aquaculture. 
    FAO Fisheries Aquaculture Technical Paper 530. Rome, Italy.

Food and Agriculture Organization (FAO). (2012). The State of World Fisheries and Aquaculture. 10 p. FAO 
    Organization of the United Nations, Rome 

Mufa’ah, & Hayati, Mardiah. (2016). Analysis Of Export Competitiveness Of Indonesian Shrimp Commodities. 
    AGRIFO Vol. 1 No.1. 

Saputri, Karunia. (2016). Opportunities and Constraints of Indonesian Shrimp Exports to the Japanese Market. 
    E- Mulawarman 5, 4, 1179-1194. 

UNCOMTRADE. (2010-2018). United Nations Commodity Trade Statistics Database – 
    get trade data (http://comtrade.un.org/db/dqQuickQuery.aspx)

Wati, L.A., Wen, Chang, and Mustadjab, M. (2013). Competitiveness of Indonesian Shrimp Compare with Thailand Shrimp in Export Market. 
    Wacana Vol. 16 No.1.
